---
title: "part_d_atc_codes_clean_csv_merge"
author: "Darya Akimova"
date: "February 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

Packages:
```{r packages}
library(tidyverse)
library(data.world)
```

Data:
```{r data}
drug.spend.ds <- "https://data.world/data4democracy/drug-spending"
# drug names from Medicare Part D data:
d.drugs <- data.world::query(
  data.world::qry_sql("SELECT DISTINCT brand_name, generic_name FROM `spending_part_d_2011to2015_tidy`"),
  dataset = drug.spend.ds
  ) %>% 
  mutate_all(str_trim)  # just in case
glimpse(d.drugs)
# atc classification system from atc_codes_clean.csv
atc <- data.world::query(
  data.world::qry_sql("SELECT * FROM atc_codes_clean"),
  dataset = drug.spend.ds
) 
glimpse(atc)
```

A bit of a mess, and too much information for what is needed right now.

```{r}
atc.sml <- atc %>%
  select(level5_code:kegg) %>% 
  arrange(level5)
glimpse(atc.sml)
map_int(map(atc.sml, unique), length)
```

It's still a bit of a mess from where I stopped in my attempt to clean up this data. Wasn't sure how the names would be used.Also note, some compounds have more than 1 level 5 code - which means that they have multiple uses according to the ATC classification.

Getting rid of the KEGG codes in the names would be a good place to start.

```{r}
atc.sml <- atc.sml %>% 
  mutate(lvl5_cln = str_replace(level5, "\\[DG.+\\]", "") %>% str_trim() %>% str_to_lower())
str_sub(atc.sml$kegg, start = 1, end = 6) <- ""
glimpse(atc.sml)
atc.sml <- atc.sml %>% 
  mutate(kegg_cln = str_replace(kegg, "\\([[:upper:]]{2,}.+\\)", "") %>% str_trim() %>% str_to_lower())
atc.sml$kegg_cln[is.na(atc.sml$kegg_cln)] <- atc.sml$lvl5_cln[is.na(atc.sml$kegg_cln)]
atc.cln <- atc.sml %>% 
  select(level5_code, kegg_cln) %>% 
  distinct() %>% 
  arrange(kegg_cln)
dim(atc.sml)
dim(atc.cln)
glimpse(atc.cln)
```

Now that that's been cleaned up a bit, let's try some matching:

```{r}
brand.semi.t1 <- d.drugs %>% 
  semi_join(atc.cln, by = c("brand_name" = "kegg_cln"))
dim(brand.semi.t1)
brand.anti.t1 <- d.drugs %>% 
  anti_join(atc.cln, by = c("brand_name" = "kegg_cln"))
dim(brand.anti.t1)
generic.semi.t1 <- brand.anti.t1 %>% 
  semi_join(atc.cln, by = c("generic_name" = "kegg_cln"))
dim(generic.semi.t1)
generic.anti.t1 <- brand.anti.t1 %>% 
  anti_join(atc.cln, by = c("generic_name" = "kegg_cln")) %>% 
  arrange(generic_name)
dim(generic.anti.t1)
matched.so.far <- brand.semi.t1 %>% 
  bind_rows(generic.semi.t1)
dim(matched.so.far)
# break 1
# write_csv(matched.so.far, "partD_matched_2034_pass1.csv")
# write_csv(generic.anti.t1, "partD_unmatched_2464_pass1.csv")
```

This is much better than the first pass with `drug_uses.csv`, where I only got 1325 matches on the first pass. A good sign! 

Now, drugs with multiple ingredients:

```{r}
generic.anti.t1 <- generic.anti.t1 %>% 
  mutate(
    alt_name = generic_name,
    mult.slash = str_count(alt_name, "/"),
    mult.comm = str_count(alt_name, ","),
    mult.dash = str_count(alt_name, "-"),
    mult.all = mult.slash + mult.comm + mult.dash
    )
glimpse(generic.anti.t1)
nrow(generic.anti.t1 %>% filter(mult.all > 0))
# from previous attempts at matching: "/pf" probably indicates a prefilled syringe
slash.pf <- generic.anti.t1 %>% filter(str_detect(alt_name, "pf"))
nrow(slash.pf)
glimpse(slash.pf)
pf.semi <- generic.anti.t1 %>% 
  mutate(alt_name = str_trim(str_replace(alt_name, "/pf", ""))) %>% 
  semi_join(atc.cln, by = c("alt_name" = "kegg_cln")) %>% 
  select(brand_name, generic_name)
dim(pf.semi)
matched.so.far <- matched.so.far %>% 
  bind_rows(pf.semi)
pf.anti <- generic.anti.t1 %>% 
  mutate(alt_name = str_trim(str_replace(alt_name, "/pf", ""))) %>% 
  anti_join(atc.cln, by = c("alt_name" = "kegg_cln"))
dim(pf.anti)
# another from previous attempts:
sum(str_detect(pf.anti$alt_name, "hcl"))
sum(str_detect(atc.cln$kegg_cln, "hcl"))
sum(str_detect(atc.cln$kegg_cln, "hydrochloride"))
# "hcl" = "hydrochloride"
pf.anti <- pf.anti %>% 
  mutate(alt_name = str_replace_all(alt_name, "hcl", "hydrochloride"))
hcl.semi <- pf.anti %>% 
  semi_join(atc.cln, by = c("alt_name" = "kegg_cln"))
dim(hcl.semi)
glimpse(hcl.semi %>% select(ends_with("name")))
# 424 more matches!
matched.so.far <- matched.so.far %>% 
  bind_rows(hcl.semi %>% select(brand_name, generic_name))
dim(matched.so.far)
dim(unique(matched.so.far))
# nice, no accidental repeats
hcl.anti <- pf.anti %>% 
  anti_join(atc.cln, by = c("alt_name" = "kegg_cln"))
dim(hcl.anti)
```

Up to 1786 of the remaining drugs may have multiple ingredients. Although that's not necessarily the case because the characters `,` and `-` especially have different uses in drug names. The `/` character is the most straightforward currently to tackle. 

```{r}
table(hcl.anti$mult.slash)
hcl.anti <- hcl.anti %>% 
  mutate(mult.slash = str_count(alt_name, "/"))
mult.ingred.5 <- hcl.anti %>% 
  filter(mult.slash >= 4) %>% 
  arrange(generic_name)
# start with 2 ingredient drugs, separated by only 1 `/`
gen.anti.2ingrd <- hcl.anti %>% 
  filter(mult.slash == 1) %>% 
  mutate(alt_copy = alt_name) %>% 
  separate(alt_copy, c("ingrd_1", "ingrd_2"), sep = "/") %>% 
  select(ends_with("name"), starts_with("ingrd"))
glimpse(gen.anti.2ingrd)
gen.anti.2ingrd <- gen.anti.2ingrd %>% 
  gather("ingred", "chem_name", 4:5) %>% 
  arrange(alt_name)
glimpse(gen.anti.2ingrd)
ingrd2.semi.t1 <- gen.anti.2ingrd %>% 
  semi_join(atc.cln, by = c("chem_name" = "kegg_cln"))
dim(ingrd2.semi.t1)
ingrd2.anti.t1 <- gen.anti.2ingrd %>% 
  anti_join(atc.cln, by = c("chem_name" = "kegg_cln"))
dim(ingrd2.anti.t1)
dim(ingrd2.anti.t1 %>% select(chem_name) %>% unique())
test <- ingrd2.anti.t1 %>% select(chem_name) %>% unique() %>% arrange(chem_name)
```

